---
title: "lazyIris examples"
author: "[Phil](https://github.com/phil8192/lazy-iris)"
date: "`r Sys.Date()`"
output: 
#  rmarkdown::html_vignette:
  html_document:
    theme: journal 
    toc: true
  pdf_document:
    toc: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Microstructure2 Guide}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
options(digits.secs=3)
library(lazyIris)
knitr::opts_chunk$set(dpi=100, fig.width=10, fig.height=6)
```
# Overview 
_lazyIris_ is a small implementation of k-nearest neighbours applied to the 
famous iris dataset.

# Installing
First, ensure that the [devtools](https://github.com/hadley/devtools) package 
is installed and then install directly from the package github 
[repository](https://github.com/phil8192/lazy-iris).

```{r}
# check for and install devtools.
if(!("devtools" %in% rownames(installed.packages())))
  install.packages("devtools")

# install and load.
if(!("lazyIris" %in% rownames(installed.packages())))
  devtools::install_github("phil8192/lazy-iris")
require(lazyIris)
```

# Loading data
The package has preprocessed iris data attached.

```{r}
attach(iris.data)
```
Example data may be loaded from the inst/extdata directory by using the 
_loadData_ function. In addition, the _checkData_ function will perform any 
necessary data sanity checks.
```{r}
iris.data <- checkData(loadData())
```

# A quick look at the data
The dataset consists of 4 features and 3 possible classes. Some of the features
are highly correlated:

```{r, results="hide"}
cor(iris.data[, 1:4])
```

```{r, echo=F, results="markup"}
knitr::kable(cor(iris.data[, 1:4]))
```

## Iris data visualisation

The package provides a means to visualise the relationship between the 4 
features and the corresponding class.

```{r}
# plot all the data.
visualise(iris.data, class.name="species", main="iris data", plot.hist=TRUE,
    plot.cor=TRUE)
```

In the above visualisation, the colours correspond to the classification of the
species of iris plant:

| colour | species          |
|-------:|:-----------------|
|    red | Iris setosa      |
|  green | Iris versicolour |
|   blue | Iris virginica   |

The lower left panels show the correlation between the 4 iris features, the
diagonal panels contain a histogram of the distribution of each feature, and
finally, the upper right panels contain scatter plots of each possible feature
combination colour coded by species.

# Querying the data

The _knn_ function makes it possible to query the data for neighbouring 
instances given an arbitrary list of features. 

## 10-Nearest neighbours

The following example obtains the _top 10_ nearest neighbours to query:

```{r, results="hide"}
# form the query.
# in this example, the feature values are actually the mean values in the
# dataset, thus the results may be interpreted as the top 10 "most average"
# instances.

query <- list(
    sepal.length=5.84,
    sepal.width=3.05,
    petal.length=3.76,
    petal.width=1.20)

# obtain the nearest-neighbours.
top.10 <- knn(query, iris.data, 10)
print(top.10, row.names=FALSE)
```

```{r, echo=F, results="markup"}
query <- list(sepal.length=5.84, sepal.width=3.05,
              petal.length=3.76, petal.width=1.20)
top.10 <- knn(query, iris.data, k=10)
knitr::kable(top.10, row.names=F)
```

## Classification

In addition to the N-nearest neighours, the function also returns the distance
from the query point. This distance can be used to predict the most likely 
class of the query point using the _classifier_ function.

```{r}
query <- list(sepal.length=5.84, sepal.width=3.05, 
              petal.length=3.76, petal.width=1.20)
top.10 <- knn(query, iris.data, k=10)
prediction <- classifier(top.10$species, top.10$distance)
print(paste("prediction =", prediction$pred,
            "confidence =", prediction$conf))
```

## Visualising a query result

Given a list of nearest neighbours returned from the _knn_ function, it is 
possible to visualise the query point and it's nearest neighbours over all 
dimensions in the feature space by using the _visualise_ function.

```{r}
# form the query and obtain the nearest neighbours.
q <- list(sepal.length=5.84, sepal.width=3.05,
          petal.length=3.76, petal.width=1.20)
top.10 <- knn(query, iris.data, k=10)

# visualise the result.
visualise(iris.data, class.name="species", query=q, neighbours=top.10,
    main="iris data neighbours", plot.hist=TRUE, plot.cor=FALSE)
```

In the above plot, the query point is shown as a black point. The resulting 
neibours from the _knn_ query are highlighted (opaque) circles. In addition, 
the query point with respect to the distribution of features has been 
highlighted with a black dashed vertical line over the corresponding feature
histograms.



